\bf Faster Neural Machine Translation Inference

### 2018-02-05 fast-beam/fast-beam.tex
fast beam paper

The output layer of most deep learning models consist of the following steps
\begin{enumerate}
   \item \vspace{-2 mm} multiplication of the input with the weight matrix $p = w x$
   \item \vspace{-2 mm} addition of a bias term to the resulting scores $p = p + b$
   \item \vspace{-2 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
   \item \vspace{-2 mm} a search for the best output class, and probability if necessary $\argmax_i p_i$
\end{enumerate}

@@

\Procedure{FindBest}{softmax vector $p$}

@@

Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from Figure~\ref{algo:Find best} to Figure~\ref{algo:Calculate softmax} while $max$ is being sought.

@@

% \begin{figure} [h]
% \begin{algorithmic}
% \REQUIRE vector $p$, bias vector $b$
% 
% \COMMENT{add bias, calculate $max$ \& $argmax$}
% 
% \STATE $max \gets - \infty$ 
% \FORALL{$p_i$ in $p$}
%   \IF{$p_i + b_i > max$}
%     \STATE $max \gets p_i + b_i$
%     \STATE $best \gets i$
%   \ENDIF
% \ENDFOR 
% 
% \COMMENT{calculate denominator}
% 
% \STATE $sum \gets 0$ 
% \FORALL{$p_i$ in $p$}
%   \IF{$p_i > max$}
%     \STATE $sum \gets sum + \exp(p_i - max)$
%   \ENDIF
% \ENDFOR 
% 
% \RETURN $\frac{1}{sum}$, $best$ 
% 
% 
% \end{algorithmic}
% \caption{Fused softmax and argmax}
% \label{algo:Fused Kernel}
% \end{figure}

In fact, we are usually only interested in the best class during inference, not the probability. Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether. %, Figure~\ref{algo:Argmax only}.

% \begin{figure} [h]
% \begin{algorithmic}
% \REQUIRE activation vector $p$, bias vector $b$
% 
% \COMMENT{add bias, calculate $argmax$}t
% \FORALL{$p_i$ in $p$}
%   \IF{$p_i + b_i > max$}
%     \STATE $max \gets p_i + b_i$
%     \STATE $argmax \gets i$
%   \ENDIF
% \ENDFOR 
% 
% 
% \RETURN $argmax$ 
% 
% \end{algorithmic}
% \caption{Argmax only}
% \label{algo:Argmax only}
% \end{figure}

@@

% \begin{figure} [h]
% \begin{algorithmic}
% %\REQUIRE source sentence $s$, translation options
% \WHILE{more input} 
%   \STATE Create batch
%   \STATE Encode
%   \WHILE{batch is not empty} 
%     \STATE Decode batch
%     \FOR{each sentence in batch}
%       \IF{translation is complete}
%         \STATE Remove sentence from batch
%       \ENDIF
%     \ENDFOR
%   \ENDWHILE
% \ENDWHILE 
% \end{algorithmic}
% \caption{Mini-batching}
% \label{algo:Mini-batching}
% \end{figure}

@@

% \begin{figure} [h]
% \begin{algorithmic}
% %\REQUIRE source sentence $s$, translation options
% \WHILE{more input} 
%   \STATE Create encoding batch
%   \STATE Encode
%   \STATE Add to queue
% \ENDWHILE 
% \end{algorithmic}
% \caption{Encoding for top-up batching}
% \label{algo:Encoding for top-up batching}
% \end{figure}

@@

% \begin{figure} [h]
% \begin{algorithmic}
% %\REQUIRE source sentence $s$, translation options
% \STATE create decoding batch $b$ from queue
% \WHILE{$b$ is not empty}
%   \STATE Decode $b$
%   \STATE Replace completed sentences with new sentence from queue
% \ENDWHILE 
% \end{algorithmic}
% \caption{Decoding for top-up batching}
% \label{algo:Decoding for top-up batching}
% \end{figure}

>>>>

The output layer of most deep learning models consist of the following steps
\begin{enumerate}
   \item \vspace{-2 mm} multiplication of the input with the weight vector $p = w x$
   \item \vspace{-2 mm} addition of a bias term to the resulting scores $p = p + b$
   \item \vspace{-2 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
   \item \vspace{-2 mm} a search for the best output class, and probability if necessary $\argmax_i p_i$
\end{enumerate}

@@

\Procedure{Find-Best}{softmax vector $p$}

@@

Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from FIND-BEST to SOFTMAX.

@@

\begin{figure} [h]
\begin{algorithmic}
\Procedure{Fused Kernel}{vector $p$, bias vector $b$}

\Comment{add bias, calculate $max$ \& $argmax$}

\State $max \gets - \infty$ 
\ForAll{$p_i$ in $p$}
  \If{$p_i + b_i > max$}
    \State $max \gets p_i + b_i$
    \State $best \gets i$
  \EndIf
\EndFor

\Comment{calculate denominator}

\State $sum \gets 0$ 
\ForAll{$p_i$ in $p$}
  \If{$p_i > max$}
    \State $sum \gets sum + \exp(p_i - max)$
  \EndIf
\EndFor

\Return $\frac{1}{sum}$, $best$

\EndProcedure
\end{algorithmic}

\caption{Fused softmax and argmax}
\label{algo:Fused Kernel}
\end{figure}

In fact, we are usually only interested in the best class during inference, not the probability. Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether, Figure~\ref{algo:Argmax only}.

\begin{figure} [h]
\begin{algorithmic}
\Procedure{Find-1-Best}{vector $p$, bias vector $b$}

\State $max \gets - \infty$ 
\ForAll{$p_i$ in $p$}
  \If{$p_i + b_i > max$}
    \State $max \gets p_i + b_i$
    \State $best \gets i$
  \EndIf
\EndFor

\Return $best$

\end{algorithmic}
\caption{Find 1 best only}
\label{algo:Argmax only}
\end{figure}

@@

\begin{figure} [h]
\begin{algorithmic}
\Procedure{Mini-Batching}{}
%\REQUIRE source sentence $s$, translation options
\While{more input} 
  \State Create batch
  \State Encode(batch)
  \While{batch is not empty} 
    \State Decode(batch)
    \For{each sentence in batch}
      \If{translation is complete}
        \State Remove sentence %from batch
      \EndIf
    \EndFor
  \EndWhile
\EndWhile

\end{algorithmic}
\caption{Mini-batching}
\label{algo:Mini-batching}
\end{figure}

@@

\begin{figure} [h]
\begin{algorithmic}

\Procedure{Encode}{}
\While{more input} 
  \State Create encoding batch $b$
  \State Encode($b$)
  \State Add $b$ to queue
\EndWhile

\end{algorithmic}
\caption{Encoding for top-up batching}
\label{algo:Encoding for top-up batching}
\end{figure}

@@

\begin{figure} [h]
\begin{algorithmic}

\Procedure{Decode}{}

\State create batch $b$ from queue
\While{$b$ is not empty}
  \State Decode($b$)
  \State Replace completed sentences with new sentence from queue
\EndWhile

\end{algorithmic}
\caption{Decoding for top-up batching}
\label{algo:Decoding for top-up batching}
\end{figure}
